---
title: "Practise 3"
author: "Maryam Ahmed "
date: "2024-10-14"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

**4.** The Capital Asset Price Model (CAPM) is a financial model that attempts to predict the 
rate of return on a financial instrument, such as a common stock, in such a way that it is
linearly related to the rate of return on the overal market. Specifically

$$
R_{StockA,i} = \beta_{0} + \beta_{1}R_{Market, i} + e_{i}   
$$

(Note: To align the notation with a stock's ``beta'', $\beta_{0} = A$ and $\beta_{1} = B$)


You are to study the relationship between the two variables and estimate the above model:

$R_{SUNCOR, i}$ - rate of return on Stock A for month $i$, $i = 1, 2, \cdots, 59$.

$R_{Market, i}$ - market rate of return for month $i$, $i = 1, 2, \cdots, 59$.


$\beta_{1}$ represent's the stocks `beta' value, or its *systematic risk*. It measure's the stocks volatility related
to the market volatility.  $\beta_{0}$ represents the risk-free interest rate.

The data appeaing in the [file](https://github.com/Statman44/Data602/blob/main/bondsdata.csv) contains the data on Suncor's rate of return and the Toronto Composite Index rate of return for 59 randomly selected months.  

Therefore $R_{SUNCOR, i}$ represents the monthly rate of return for a common share of Suncor stock; 
$R_{TSE, i}$ represents the monthly rate of return (increase or decrease) of the TSE Index for the same month, month $i$.
The first column in this data file contains the monthly rate of return on Suncor stock; the second column contains the monthly rate of return on the TSE index for the *same month*. 

Read this data into R Studio and answer the questions posed below. 
```{r}
library(ggplot2)
```


```{r}
# Read the CSV file from the URL
capmdata.df <- read.csv("https://raw.githubusercontent.com/Statman44/Data602/main/capm.csv")

# Check the first three rows of the data frame
head(capmdata.df, 3)  # Corrected variable name

```

----------------------------------------------------------------------------------------------------------------------------------

(a) Appropriately visualize these data. What can you infer from this visualization? Provide a brief commentary.


```{r}
plot(capmdata.df$TSE.Index,capmdata.df$Suncor,xlab="Scatter Plot of TSE Index vs. Suncor Return ",
     ylab="TSE Return",main ="Suncor Return")
```

**Answer**
It does look like Return of Suncor stock and the TSE index are positively correlated, because it does follow a positive line as you can see a trend upwards, but because of how far apart the values are from eachother linearly it does look very weakly correlated.


----------------------------------------------------------------------------------------------------------------------------------

(b) Estimate the model above.  

```{r}
model = lm(Suncor ~ TSE.Index, data = capmdata.df)
model 
```

**Answer**

model = 0.01665 +  0.53869  * (Return TSE Index)



----------------------------------------------------------------------------------------------------------------------------------


(c) In the context of these data, interpret the meaning of your estimates of the estimates $\widehat{\beta}_{0}$ and $\widehat{\beta}_{1}$, in the context of these data.

**Answer**
The model tells us that for every one-unit increase in the TSE Index, the Suncor return is expected to increase by 0.53869, assuming all other factors remain constant. Additionally, when the TSE Index is zero, the model b0 says the Suncor return will be 0.01665. This constant term represents the baseline level of the Suncor return when the TSE Index does not contribute to the prediction or better yet when it is 0. 



----------------------------------------------------------------------------------------------------------------------------------



(d) Refer to your answer in (b) In a certain month, the rate of return on the TSE Index was 4%. Predict the rate of return on Suncor stock for the same month. 

```{r}
0.01665 +  (0.53869*0.04)
```

**Answer** 

0.0381976 = 3.82%


----------------------------------------------------------------------------------------------------------------------------------


(e) Think about the conditions of this model *in the context* of these data. Create the visualizations that inspect each of the two conditions and provide commentary that addresses the validity (or invalidity) of each. 

**Answer**

There's a few conditions in this model: 

1) Confirm that the residuals from the model, the difference between actual Suncor return and predicted turn based on TSE index follows a normal distribution

2) Given that they have a linear relationship, we need to check that the residuals maintain a constant variance across different levels of the TSE index. This can be done by examining whether there is a pattern in the fitted values and residual plots.

To check for normality in residuals...

```{r}
ggplot(data = capmdata.df, aes(sample = residuals(model))) +
  stat_qq(size = 2, color = "blue") +
  geom_qq_line(color = "red") +
  ggtitle("Normal Distribution Plot of Residuals") +
  theme_minimal()
```
**Answer** We can conclude that the the residuals are slightly normal given the QQ plot of residual values follow a close normal line. 


To check for relationship...

```{r}

plot(fitted(model), resid(model),
     main = "Residuals vs Fitted Values",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 20, col ="black")
abline(h = 0, col = "red", lwd = 2, lty = 2)

```
**Answer**

Upon examining the graph, we can conclude that it appears to exhibit heteroskedasticity, as there is a discernible pattern suggesting clustering primarily around the positive fitted values. In the absence of such a pattern, we would expect the points to be scattered throughout the graph. Therefore, I assert that there is no linear relationship present, indicating that the second condition is not satisfied. However, with more data points, we may achieve a clearer graph, and the assessment could differ.


----------------------------------------------------------------------------------------------------------------------------------


(f) From these data, can you infer that the monthly rate of return of Suncor stock can be expressed as a positive linear function of the monthly rate of return of the TSE Index? State your statistical hypotheses, compute (and report) both the test statistic and the $P$-value and provide your decision. 
Ho = B1 = 0 
Ha = B1 > 0

```{r}
summary(lm(capmdata.df$Suncor ~ capmdata.df$TSE.Index))
```

```{r}
#t-value = 2.809

1 - pt(2.809, model$df)

```
**Answer** 

Given the p-value of 0.0034 and is less than 0.05, we can confirm that we would reject to the null in favor of the alternative in which we can say that that it does have a positive linear relationship.


----------------------------------------------------------------------------------------------------------------------------------


(g) Compute a 95% confidence interval for $\beta_{1}$, then interpret its meaning in the context of these data. 


```{r}
# B1 +- t* SE(B1)

0.53869 - qt(0.975, df = 57)*0.19178 
0.53869 + qt(0.975, df = 57)*0.19178 

#or 

confint(model, level = 0.95)
```
**Answer** 

B1 or TSE index CI = [ 0.15, 0.92] with 95% confidence. Which means that we are 95% confident that the true value of B1 or TSE index lies between 0.15 and 0.92. Telling us that for each 1 unit increase in suncor return, holding all other values constant, would be an increase between 0.15 and 0.92.


----------------------------------------------------------------------------------------------------------------------------------

(h) Compute a 95% confidence interval for the mean monthly rate of return of Suncor stock when the TSE has a monthly rate of return of 3%. 

```{r}
model = lm(Suncor ~ TSE.Index, data = capmdata.df)

predict(model, data.frame(TSE.Index=0.03),interval="conf", level = 0.95)
```
**Answer** 

B1 or TSE index when the value is at 3%, the monthly rate of return for Suncor will be between [0.0077 and 0.05796] at 95% confidence. 


----------------------------------------------------------------------------------------------------------------------------------

(i) In a month of September, the TSE Index had a rate of return of 1.16%. With 95% confidence, compute the September rate of return for Suncor stock. 
```{r}

predict(model, newdata=data.frame(TSE.Index=0.0116),interval='prediction', level = 0.95)

```
**Answer** 
B1 or TSE index when the value was at 1.16%, the monthly rate of return for Suncor in Septembe would have between [-0.1587 and 0.20456] at 95% confidence. 


----------------------------------------------------------------------------------------------------------------------------------



(j) Recall the Bootstrap Method. From these data, use the bootstrap method to create a 95% confidence interval for mean monthly rate of return of Suncor stock when the TSE has a monthly rate of return of 3%. Compare your result to your result in part (h). Use 1000 iterations for your bootstrap. *Carefully* consider how you would resample bivariate data points $(x_{TSE, i}, y_{Suncor, i})$. 

```{r}
set.seed(1)
nsims = 1000
TSE.Val = 0.03
predict = numeric(1000)
x_y = dim(capmdata.df)[1]

for(i in 1:nsims)
{
  index = sample(x_y,replace=TRUE)
  dsample = capmdata.df[index, ]
  predictivemodel = lm(Suncor ~ TSE.Index, data = dsample)
  predict[i] = (coef(predictivemodel)[1]) + ((coef(predictivemodel)[2])*TSE.Val)
}

bootstrap = data.frame(predict)
head(bootstrap,10)

```
```{r}
quantile(bootstrap$predict, probs = c(0.025, 0.975))
```
**Answer**

Looking at the bootstrap confidence interval, I observe that the 95% confidence interval for Suncor's return is quite similar to the interval obtained previously. Using the bootstrap method, the CI is [0.0094, 0.05695], while the prediction CI from part h was [0.0077, 0.05796]. The bootstrap interval is slightly narrower, which might suggest a more precise estimate of the true return at a 95% confidence level.
